rang

### 2023-10-30 paper/gesis_2023/index.qmd
Add gesis slides

---
title: "rang"
subtitle: "(Re)constructing reproducible R computational environment"
author: Chung-hong Chan, David Schoch
institute: GESIS
date: "2023-07-21"
date-format: "YYYY-MM-DD"
footer: "https://chainsawriot.github.io/basel2023_rang"
embed-resources: true
format: fakegesis-revealjs
---
## Computational Reproducibility

A result is reproducible when the same analysis steps performed on the same dataset **consistently produce the same answer**. This reproducibility of the result can be checked by the original investigators and other researchers within a local environment.

::: {style="font-size: 0.3em"}
Schoch, Chan, Wagner, & Bleier. (2023) Computational Reproducibility in Computational Social Science. [arXiv](https://doi.org/10.48550/arXiv.2307.01918)
:::

## Case study I: dplyr

```{r}
#| echo: true
data <- data.frame(id = c(1,1,2), outcome = c(TRUE, TRUE, FALSE))
data
```

## Case study I: dplyr

@@

## Pit stop 1: Summary

1. Computational reproducibility is tacitly dependent on the computational environment
2. Recursive dependencies can best be tackled with time-based pinning or lockfile-based system
3. P3M has its limit (2017-10-10).

::: {.notes}
P3M also has an inconvenience limitation: What if it's no longer available (like MRAN before it)?
:::

## Case study II

![](comps.png)

::: {style="font-size: 0.3em"}
Source: @whydoesr@botsin.space
:::

## Case study II

```r
unix.time(mean(rnorm(100)))
n _ 10
require(sna)
x _ rgraph(n, mode = "graph")
evcent(x)
```

## On my own machine: R 4.3.1

```{r}
#| echo: true
#| error: true
unix.time(mean(rnorm(100)))
n _ 10
suppressPackageStartupMessages(require(sna))
x _ rgraph(n, mode = "graph")
evcent(x)
```

Explanations:

```r
?`base-defunct`
```

## Non-method #1: P3M/renv

It's base. You can't pin `unix.time()` with P3M/`renv`.

::: {.notes}
in Norm Matloff's voice
:::

## Method #2?: Docker / Rocker

```bash
# work
docker run --rm -it rocker/r-ver:3.0.0
```

```text
> unix.time(mean(rnorm(100)))
   user  system elapsed 
      0       0       0 
> n _ 10
Error: unexpected input in "n _"
> 
```

```bash
# not work
docker run --rm -it rocker/r-ver:2.15.3
```

::: {.notes}
You probably will have problem installing sna
:::

## Pit stop 2: Summary

1. Computational reproducibility is tacitly dependent on the computational environment
2. It's not always about packages, the programming language changes too
3. Rocker has its limit (R 3.0.0)

::: {.notes}
(think Python 2 -> 3)

Rocker: Many old images are not usable anymore (either hardcoded MRAN or expiring Debian)
:::

@@

![](rang_logo.png){width=300,fig-align="left"}

```r
install.packages("rang")
## remotes::install_github("chainsawriot/rang")
```

::: {style="font-size: 0.3em"}
Chan & Schoch (2023). rang: Reconstructing reproducible R computational environments. PLOS One [doi](https://doi.org/10.1371/journal.pone.0286761)
:::

## Features (CRAN version)

@@

## README

```text
This README offers some guidance on how to work with the included docker container 
which was created with the R package rang on 2023-07-17

# Installing docker

The installation depends on the OS you are running
...
```

@@

## Full bash

```bash
Rscript -e "rang::resolve(\"sna\", snapshot_date = \"2002-02-01\") |> 
rang::dockerize(\"~/scoms\", cache = TRUE)"
cd ~/scoms
docker build -t scoms .
docker run --rm --name "scoms" -ti scoms
```
## For compatibility with S

![](base2.png)

## Many features not demo.

* More examples in [our paper](https://doi.org/10.1371/journal.pone.0286761)

Upcoming features

* Apptainer (née Singularity) - Thanks to the contribution from Egor Kotov
* Research Compendium
* Post installation steps, e.g. installing Quarto

## Summary

## Summary

:::: {.columns}

:::{.column width="80%"}
1. Computational reproducibility is tacitly dependent on the computational environment
2. Use P3M / renv for projects after 2017-10-10
    * only at the level of R packages
3. Plus Rocker for projects newer than R 3.0.0
4. Use `rang` for probably all projects
:::

:::{.column width="20%"}
![](rang_logo.png){width=300}
:::

::::

::: {style="font-size: 0.3em"}
Schoch, Chan, Wagner, & Bleier. (2023) Computational Reproducibility in Computational Social Science. [arXiv](https://doi.org/10.48550/arXiv.2307.01918)

Chan & Schoch (2023). rang: Reconstructing reproducible R computational environments. PLOS One [doi](https://doi.org/10.1371/journal.pone.0286761)
:::

>>>>

---
title: "rang"
subtitle: "(Re)constructing reproducible R computational environment"
author: Chung-hong Chan, David Schoch
institute: GESIS
date: "2023-11-02"
date-format: "YYYY-MM-DD"
footer: "rang"
embed-resources: true
format: fakegesis-revealjs
---

## Case study

@@

@@

@@

@@

# Summary

* code edit
* Install old version
* renv
* P3M
* rang
* New development: evercran (CRAN snapshot and versioned Docker images)
